MLH1 Results Outline

### 2020-02-16 doc/Results.rmd
adding notes and cutting fat for Q1-SC results

```{r MixedModel.2, echo=FALSE,  digits = 3}
#strain = fixed
glm.strain.fixed.full <- glm(mean_co ~ subsp * sex * strain, data=DF.HetC.MixedModel.HQ)

@@

```{r MixedModel.3, echo=FALSE, digits = 3}

@@

```{r Q1.Short.Long_biv.setup, echo=FALSE}
Q1.real.short_mouse_table <- ddply(real.short.bivData, c("mouse"), summarise,
                        n_obs = length(Obj.ID),
                         mean_SC = mean(chromosomeLength)
                         )

Q1.real.short_mouse_table <- add_strain(Q1.real.short_mouse_table)
Q1.real.short_mouse_table <- add_sex(Q1.real.short_mouse_table)
Q1.real.short_mouse_table <- add_subsp(Q1.real.short_mouse_table)

@@

#MSM
#Q1_mus.av_shortSC_t.test.MSM <- t.test((Q1.real.short_mouse_table %>% filter(sex #== "male") %>% filter(strain == "MSM") )$mean_SC,
#                           (Q1.real.short_mouse_table %>% filter(sex == "female") #%>% filter(strain == "MSM") )$mean_SC)

the pattern is generally held for comparisons using the reduced single bivalent data sets (short bivlents are the most comparable set, sinceit is unlikely to include the XX in the female data

(t.test; p= `r Q1_mus.av_shortSC_t.test.PWD$p.value`, `r Q1_mus.av_shortSC_t.test.KAZ$p.value`, and `r Q1_mus.av_shortSC_t.test.SKIVE$p.value`

for PWD, KAZ, SKIVE and MSM.

```{r Q1_SC.biv.long_ttest, echo=FALSE}

Q1.SC.short.mouse_scatter_boxplot
```

```{r show.long.biv, echo=FALSE}
#mouse av
Q1.SC.long.mouse_scatter_boxplot <- ggplot( Q1.real.long_mouse_table, aes(y = mean_SC, x = sex, color=strain)) + geom_jitter()+ geom_boxplot(alpha=.2) +scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales="free")+ylim(c(75,150) )+ggtitle("Mouse average SC Length \n long bivalent dataset")

Q1.SC.long.mouse_scatter_boxplot
```

mixed models and glms were donw to test for evolution

@@

Q1.lmer_totSC.mouse.av.table <- totSC.mouse.av.table %>% filter(strain != "AST" & strain != "TOM" & strain != "CZECH" & strain != "MOLF")
str(Q1.lmer_totSC.mouse.av.table)

@@

```{r Q1.glm_totalSC, echo=FALSE}
#total-SC

@@

@@

```{r Q1.SC_short_t.test, echo=FALSE}

```{r Q1_short_biv_lmer, echo=FALSE}
#Mouse.Table_BivData_4MM #this one is the HetC table (all the Curated BivData)
#Long.biv_mouse_table_HetC  (long biv data set for HetC models)

These plot above show that qualitatively, the average female bivalent SC length is significantly longer than the mean length for males from the same strain.

@@

$$mouse \ average \ SC  \ length ~=~ sex * strain + \varepsilon $$

```{r Q1.SC.Length_M2, echo=FALSE}
#feed these 2 DFs into M2
#Mouse.Table_BivData_4MM #this one is the HetC table (all the Curated BivData)
#Long.biv_mouse_table_HetC  (long biv data set for HetC models)
#glm(metric ~ strain)  ---mean_SC

Q1.glm_SCLength_M2 <- glm(mean_SC ~ strain * sex, data=Mouse.Table_BivData_4MM)
#summary(Q1.glm_SCLength_M2)

#Dom.all
Q1glm_SCLength_M2.Dom <- glm(mean_SC ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Dom",])
#summary(Q1glm_SCLength_M2.Dom)

#Dom.1CO
Q1glm_SCLength_M2.1CO.Dom <- glm(mean.SC_1CO ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Dom",])
#summary(Q1glm_SCLength_M2.1CO.Dom)

#Dom.2CO
Q1glm_SCLength_M2.2CO.Dom <- glm(mean.SC_2CO ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Dom",])
#summary(Q1glm_SCLength_M2.2CO.Dom)

#Musc.all
Q1glm_SCLength_M2.Musc <- glm(mean_SC ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Musc",])
#summary(Q1glm_SCLength_M2.Musc)

#Musc.1CO
Q1glm_SCLength_M2.1CO.Musc <- glm(mean.SC_1CO ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Musc",])
#summary(Q1glm_SCLength_M2.1CO.Musc)

#Musc.2CO
Q1glm_SCLength_M2.2CO.Musc <- glm(mean.SC_2CO ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Musc",])
#summary(Q1glm_SCLength_M2.2CO.Musc)

#Mol.all
Q1glm_SCLength_M2.Mol <- glm(mean_SC ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Mol",])
#summary(Q1glm_SCLength_M2.Mol)

#Mol.1CO
Q1glm_SCLength_M2.1CO.Mol <- glm(mean.SC_1CO ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Mol",])
#summary(Q1glm_SCLength_M2.1CO.Mol)

#Mol.2CO
Q1glm_SCLength_M2.2CO.Mol <- glm(mean.SC_2CO ~ strain * sex, data=Mouse.Table_BivData_4MM[Mouse.Table_BivData_4MM$subsp == "Mol",])
#summary(Q1glm_SCLength_M2.2CO.Mol)
```

@@

```{r Q1.SC.M2plot, echo=FALSE}

@@

```{r Q1.SC.long.biv_wacko, echo=FALSE}
#removed these plots for space

@@

<!-- 
lmr - sex is most significant effect

plots,

- ranges overlap,

- try the short and long mouse averages

G and MSM female?

G male vs female

>>>>

```{r MLH1.glmM2, echo=FALSE,  digits = 3}
#strain = fixed
glm.strain.fixed.full <- glm(mean_co ~ subsp * sex * strain, data=DF.HetC.MixedModel.HQ)

@@

```{r MLH1.glm.M3, echo=FALSE, digits = 3}

@@

```{r Q1.Short.Long_biv.setup, echo=FALSE}

Q1.real.short_mouse_table <- ddply(real.short.bivData, c("mouse"), summarise,
                        n_obs = length(Obj.ID),
                         mean_SC = mean(chromosomeLength)
                         )

Q1.real.short_mouse_table <- add_strain(Q1.real.short_mouse_table)
Q1.real.short_mouse_table <- add_sex(Q1.real.short_mouse_table)
Q1.real.short_mouse_table <- add_subsp(Q1.real.short_mouse_table)
Q1.real.short_mouse_table <- add_species(Q1.real.short_mouse_table)

Q1.real.short_mouse_table$subsp <- factor(Q1.real.short_mouse_table$subsp, ordered = FALSE)
Q1.real.short_mouse_table$sex <- factor(Q1.real.short_mouse_table$sex, ordered = FALSE)
Q1.real.short_mouse_table$strain <- factor(Q1.real.short_mouse_table$strain, ordered = FALSE)
Q1.real.short_mouse_table$species <- factor(Q1.real.short_mouse_table$species, ordered = FALSE)

@@

@@

Q1.lmer_totSC.mouse.av.table <- totSC.mouse.av.table %>% filter(strain != "AST" & strain != "TOM" & strain != "CZECH" & strain != "MOLF")

#str(Q1.lmer_totSC.mouse.av.table)

@@

```{r Q1.glm_totalSC, echo=FALSE, }
#total-SC

@@

Q1.SC.short.mouse_scatter_boxplot

#ploting SKIVE alone, confirms the ranges overlap, KAZ also have close ranges
skiveQ1.SC.short.mouse_scatter_boxplot <- ggplot( Q1.real.short_mouse_table %>% filter(strain == "SKIVE"), aes(y = mean_SC, x = sex, color=strain)) + geom_jitter()+ geom_boxplot(alpha=.2) +scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales="free")+ylim(c(20,115) )+ggtitle("Mouse average SC Length \n short bivalent dataset")
#skiveQ1.SC.short.mouse_scatter_boxplot
#there are 2 lowish female means,

When skive is plotted alone it does have overlapping ranges.

#MSM
#Q1_mus.av_shortSC_t.test.MSM <-t.test( (Q1.real.short_mouse_table %>% filter(sex == "male") %>% filter(strain == "MSM") )$mean_SC, 
#          (Q1.real.short_mouse_table %>% filter(sex == "female") %>% filter(strain == #"MSM"))$mean_SC )
#1 male, 3 females measures
#try to merge the manual - whole cell observations (at least 2 males)

The pattern is generally held for comparisons using the reduced single bivalent data sets (short bivlents are the most comparable set, sinceit is unlikely to include the XX in the female data

(t.test; p= `r Q1_mus.av_shortSC_t.test.PWD$p.value`, `r Q1_mus.av_shortSC_t.test.KAZ$p.value`, and `r Q1_mus.av_shortSC_t.test.SKIVE$p.value` for PWD, KAZ, SKIVE

for the short bivalents are not significantly differen, for this strain there are 4 female and 4 male

** there are more SKIVE female mouse observations -- but they are filtered out because I excluded them from the MLH1 dataset.  -- for low quality -- they might have good total SC spreads -- // but

* it think total sc only has 1 female observation but the short bivalent has 4 skive females

Mixed models for the sex differences in mouse averages for short bivalents (reduced dataset)

$$mouse \ average \ short \ SC  \ length ~=~ subsp * sex + rand(strain) + \varepsilon $$

```{r Q1_short_biv_lmer, echo=FALSE}

lmer.Q1_SCLength_short_M1 <- lmer(mean_SC ~ subsp * sex + (1|strain), data=Q1.real.short_mouse_table, REML = TRUE)

lmer.Q1_SCLength_short_M1_R.sex <- lmer(mean_SC ~ subsp + (1|strain), data=Q1.real.short_mouse_table, REML = TRUE)

lmer.Q1_SCLength_short_M1_R.subsp <- lmer(mean_SC ~ sex + (1|strain), data=Q1.real.short_mouse_table, REML = TRUE)

Sex is the most significant factor for the mixed model of mean short bivalents.

$$mouse \ average \ short \ SC  \ length ~=~ subsp * sex * strain + \varepsilon $$

```{r Q1.SC.Length_M2, echo=FALSE}

Q1.real.short_mouse_table$subsp <- factor(Q1.real.short_mouse_table$subsp, ordered = FALSE)
Q1.real.short_mouse_table$sex <- factor(Q1.real.short_mouse_table$sex, ordered = FALSE)
Q1.real.short_mouse_table$strain <- factor(Q1.real.short_mouse_table$strain, ordered = FALSE)

Q1.glm_short.SC_M2 <- glm(mean_SC ~ subsp * strain * sex, data=Q1.real.short_mouse_table)

#summary(Q1.glm_short.SC_M2)

summ.Q1.glm_short.SC_M2 <- summary(Q1.glm_short.SC_M2)

round(summ.Q1.glm_short.SC_M2$coefficients[,c(1,4)],5)

#skive strin very slight sign
#male highly sig
#skive * male slightly sig

skive strin very slight sign, male highly sig,  skive * male slightly sig

$$mouse \ average \ short \ SC  \ length ~=~ sex * strain + \varepsilon $$

```{r Q1.glm_short.SC_M3, echo=FALSE}

Q1.glm_short.SC_M3 <- glm(mean_SC ~ strain * sex, data=Q1.real.short_mouse_table)

summary(Q1.glm_short.SC_M3)

summ.Q1.glm_short.SC_M3 <- summary(Q1.glm_short.SC_M3)

round(summ.Q1.glm_short.SC_M3$coefficients[,c(1,4)],5)
#sex male very sig -- only sig
```

male is the only significant p value from this model.

```{r show.long.biv, echo=FALSE, include=FALSE}
#mouse av
Q1.SC.long.mouse_scatter_boxplot <- ggplot( Q1.real.long_mouse_table, aes(y = mean_SC, x = sex, color=strain)) + geom_jitter()+ geom_boxplot(alpha=.2) +scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales="free")+ylim(c(75,150) )+ggtitle("Mouse average SC Length \n long bivalent dataset")

```{r JUNK.Q1.SC_short_t.test, echo=FALSE, include=FALSE}

```{r Q1.strain.specific.junk, include=FALSE}

@@

```{r Q1_SC.biv.long_ttest, echo=FALSE, include=FALSE}

@@

@@

```{r Q1.SC.M2plot, echo=FALSE, include=FALSE}

@@

```{r Q1.SC.long.biv_wacko, echo=FALSE, include=FALSE}
#removed these plots for space

@@
